---
title: "test"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)

```

# Load Data

```{r}

# The data was provided by Fiona Tummon, in form of 10-minute pollen counts.
# The data prep until that point was not re-evaluated.
# The datetimes were provided in a seperate file (tbd whether those are UTC or not)
# PAY2, PAY4 and PAY8 are three different pollen traps
# lines 2,4,6,7 are four different lines, along which the pollen were manually counted under the microsope.
# Lines 2 and 6 were counted in 2013, whereas lines 4 and 7 were counted in 2019

path_data <- paste0(here::here(), "/ext-data/")
files_data <- list.files(path_data, pattern = "2013_10min.csv")
files_datetimes <- list.files(path_data, pattern = "dates_10min.csv")

data_raw <-
  map(files_data, ~ read_delim(
    paste0(path_data, .x),
    delim = ";",
    col_names = TRUE,
    col_types = cols(.default = "n"),
    na = "nan"
    )
  ) %>% 
  setNames(files_data)

dates_raw <-
  map(files_datetimes,~ read_delim(
    paste0(path_data, .x),
    delim = ";",
    col_names = FALSE,
    col_types = cols(.default = "n") # dates and times are in a weird format, have to import as numeric as a consequence
    )
  ) %>% 
  setNames(files_datetimes)

dates <- map(dates_raw, ~mutate(
  .x,
  time = sprintf("%04d", as.integer(X2)), # otherwise we lose leading zeros
  datetime = ymd_hm(paste0(X1, time)))%>% 
    select(datetime)
  )

# Combining the measurements with the datetimes for the three traps
# It appears that the traps have slightly different timestamps
data <- map2(data_raw, rep(1:3, each = 4), ~bind_cols(.x, dates[[.y]]) %>% mutate(trap = 2^(.y)))

# Create a new variable to differentiate between counted lines; note that traps were already defined in the previous map-call above
data <- map2(data, rep(c(2, 4, 6, 7), 3), ~mutate(.x, line = .y))

# We are only interested on a subset of the data
# The rder of variables is different in the PAY2 and PAY6 data, those are the old counts from 2013
species_selection <- c("Castanea", "Alnus", "Ulmus", "Cupress", "Fraxinus", "Fagus", "Juglans", "Plantago", "Corylus", 
    "Pinus", "Quercus", "Rumex", "Platanus", "Populus", "Poaceae", "Salix", "Betula", "Carpinus", 
    "Urticacées", "Taxus", "Picea")

data = map(data, ~select(.x, 
    kacastz0, kaalnuz0, kaulmuz0, kacuprz0, kafraxz0, kafaguz0, kajuglz0, khplanz0, kacoryz0, 
    kapinuz0, kaquerz0, khrumez0, kaplatz0, kapopuz0, khpoacz0, kasaliz0, kabetuz0, kacarpz0, 
    khurtiz0, kataxuz0, kapicez0, datetime, trap, line) %>% 
  setNames(c(species_selection, "datetime", "trap", "line")))

# map(dates, ~summary(.x))
# map(data_raw, ~summary(.x))


```

There are a few missing measurements, we will impute/exclude those in the next step.
Generally, there if one species is NA at a certain time, then all species are NA.
Hence we could assume that the traps had a malfunction during that time.
What is surprising though is the fact that not all lines have the same amount of NAs.
This might suggest that some entries were omitted during the manual pollen counting exercise.
If these values are outside of the blooming season then we don't have an issue, I will check this below.

```{r}

map(data, ~.x %>% filter(is.na(Populus)) %>% nrow)

```

The team in Payerne fuond that the air sucking rates of the Hirst traps are actually higher in reality than reported by the manufacturer.
The traps suck in 13.5 l per minute instead of 10 l per minute. Hence our Pollencounts per 10minutes are too high and should be reduced by a factor of 1.35.

```{r}
data <- map(data, ~mutate_at(.x, vars(species), ~./1.35))
```


Now let's plot some time series to get a better overview of the data.
For better visibility I am displaying hourly sums as a minimum timespan.

```{r}

species <- "Ulmus"

data[[1]] %>% 
  ggplot(aes(x = datetime)) +
  geom_point(aes_string(y = species))

data[[1]] %>% 
  # filter(!!sym(species) != 0) %>% 
  ggplot() +
  geom_boxplot(aes_string(y = species))

data[[1]] %>% select(Poaceae) %>% unique

```












